---
title: "Some practice"
format: 
  html:
    self-contained: true
    code-line-numbers: true
  pdf: 
    documentclass: article
custom_title_page: false
filters:
  - lightbox
lightbox: auto
---

---
subtitle: "Let's practice some usefull commands"
date: 2023-02-27

author:
  - name: Javier Silva-Valencia
    orcid: 0000-0002-5982-2821
    email: javier.silva@unmsm.edu.pe
    affiliations:
      - name: Instituut Voor Tropische Geneeskunde. Antwerp-Belgium

language: 
  title-block-author-single: "Writer"
  
abstract: Today we had an introduction to how to create subsets by row, subset by colls, appending (joining up-down), merging (joining left-righ).
---

------------------------------------------------------------------------

------------------------------------------------------------------------

### Exercise 1

Let's practice dividing and joining data.

#### Import CSV

Importing the CSV database under the name of Q1_B, with "," as separator, and with "." as decimal:

```{r}
Q1_B <- read.csv("C:/Users/pined/OneDrive - Universidad Nacional Mayor de San Marcos/Javier 2022/Belgica/AC2_DataAnalysis_ThWk/Material/Q1_B.csv", sep=",", dec= ".")
```

#### Create a subset by row

Checking the identification variable of this dataset (FSN)

```{r}
summary(Q1_B$FSN)
```

::: callout-tip
The median is 52032
:::

Let's divide the data in two parts taking as cut point the half of the rows (The median of FSN)

```{r}
top <- subset(Q1_B, subset= FSN < 52032)
bottom <- subset(Q1_B, subset= FSN >= 52032)
```

Seeing the structure of the original dataset and the new subsets

```{r}
str(Q1_B)     #Original dataset
str(top)      #Subset top of the data
str(bottom)   #Subset bottom of the data
```

::: callout-tip
In Q1_B we have 13377 obs

If we sum the observations in the subsets:

Top + bottom: 6688 + 6689 =13377

So, it is ok
:::

#### Joining (Append)
Let's join the two subsets (top and bottom)

```{r}
merged_top_bottom  <-  rbind(top,bottom)  #To merge
str(merged_top_bottom)                    #To see the result
```

::: callout-tip
There are 13377 obs 

It is the same of the original, So it is ok
:::

#### Create a subset by coll

Seeing the names of the collums (name of the variables)

```{r}
names(Q1_B)
```

Let's divide the data in two parts taking as cut point the half of the columns

We have to be careful to keep an identification variable in each subset (in this case is FSN)

```{r}
Left <- subset(Q1_B, select=c("ID", "FSN", "Neem_Tree", "Neem_Tree_Distance", "Neem_Tree_Size", "Neem_Tree_Age", "Neem_Tree_Usage", "Neem_Tree_Use_Other", "Bamboo_Tree", "Bamboo_Tree_Dist", "Banana_Tree", "Banana_Tree_Dist", "Rice_Field", "Rice_Field_Dist", "Perm_Water_Body", "Perm_Wat_Body_Dist",  "Wat_Body_Mid_Point", "No_Mosquito_Net", "Sprayed_2010", "Sprayed_2009", "Floor_Material"))
str(Left)
```

```{r}
Rigth <- subset(Q1_B, select=c("FSN", "Other_Floor_Mat", "Is_Floor_Damp", "Roof_Material", "Other_Roof_Material", "Wall_Material", "Other_Wall_Material", "Windows_in_Room", "Granaries_in_HH", "Source_Drink_Water", "Other_Src_Drink_Wat", "Toilet_Facility", "Other_Toilet_Fac", "Cooking_Fuel", "Other_Cooking_Fuel", "Source_Light", "Other_Source_Light", "Electricity_in_HH", "No_Of_Rooms", "No_Sleeping_Rooms"))

#Or select the opposite (With a "-" in front of the c):
#Rigth <- subset(Q1_B, select=-c("ID", "Neem_Tree", "Neem_Tree_Distance", "Neem_Tree_Size", "Neem_Tree_Age", "Neem_Tree_Usage", "Neem_Tree_Use_Other", "Bamboo_Tree", "Bamboo_Tree_Dist", "Banana_Tree", "Banana_Tree_Dist", "Rice_Field", "Rice_Field_Dist", "Perm_Water_Body", "Perm_Wat_Body_Dist",  "Wat_Body_Mid_Point", "No_Mosquito_Net", "Sprayed_2010", "Sprayed_2009", "Floor_Material"))

str(Rigth)
```

#### Joining (Merge)

Let's join the two subsets (left and right) according to the FSN 

```{r}
merged_left_right  <-  merge(Left, Rigth, all=TRUE, by = "FSN")
str(merged_left_right)
```



### Exercise 2

Let's practice renaming variables and replacing values

```{r}
Q1_B_1 <- read.csv("C:/Users/pined/OneDrive - Universidad Nacional Mayor de San Marcos/Javier 2022/Belgica/AC2_DataAnalysis_ThWk/Material/Q1_B_1.csv", sep=",", dec= ".")
```

```{r}
str(Q1_B_1)
head(Q1_B_1)
```

::: callout-tip
We observe: 

- There are 9029 obs 

- The animal name is quite large sometimes 

- The variable names are quite large
:::


#### Renaming variables

Renaming variables labels: In collumns 2-5 to "anim", "count", "dist", "indor", "daysin"

```{r}
names(Q1_B_1)[c(2,3,4,5,6)] <- c("anim", "count", "dist", "indor", "daysin")
str(Q1_B_1)
#View(Q1_B_1)
```

```{r}
table(Q1_B_1$count)
table(Q1_B_1$dist)
table(Q1_B_1$indor)
table(Q1_B_1$daysin)
```

#### Replacing values in observations

```{r}
table(Q1_B_1$anim)
Q1_B_1$anim[Q1_B_1$anim =="Buffaloes"] <- "Buff"
Q1_B_1$anim[Q1_B_1$anim =="Goats"] <- "Goat"

#But if we want to keep the 4 (x) first characters in a all rows
Q1_B_1$anim <- substr(Q1_B_1$anim, 0, 3)
table(Q1_B_1$anim)
str(Q1_B_1)
```


We are going to save this data because we are going to use it later
```{r}
saveRDS(Q1_B_1, "Q1_B_1.RDS")
```
